Lobbying and Legislative Uncertainty

### 2016-04-11 presentation.tex
remove stuff I REALLY don't need from presentation

\title[Endogenous Politics and the Design of Trade Institutions\hspace{2.05in}\insertframenumber/\inserttotalframenumber]{Endogenous Politics and \\ the Design of Trade Institutions}
%\author[Kristy Buzard]{Kristy Buzard \\ Syracuse University and The Wallis Institute \\ kbuzard@syr.edu}
\author[Kristy Buzard]{\texorpdfstring{Kristy Buzard\newline Syracuse University and The Wallis Institute  \newline\url{kbuzard@syr.edu}}{Kristy Buzard}}
\date{April 6, 2016}
\begin{document}
\maketitle
%\insertpresentationendpage removed b/c of appendix

@@

\begin{frame}{Role of Trade Agreements: TOT Externality}
\pause
Grossman and Helpman (1995)
\pause
\begin{itemize}[<+->]
	\item Add endogenous politics
	\item Now in ``Trade War'': two reasons for positive tariff
		\begin{itemize}
			\item TOT externality + pressure from import competing lobby
		\end{itemize}

	\item Trade agreement: only internalizes TOT externality
\end{itemize}
\end{frame}

\begin{frame}{Role of Trade Agreements: New Trade Theory}
\pause
``New Trade Theory'' externalities
\pause
\begin{itemize}[<+->]
	\item Source: imperfect competition (firm level)
		\begin{itemize}
			\item Delocation: Ossa 2011, Bagwell and Staiger 2012
			\item Profit Shifting: Mrazova 2011, Ossa 2012  
		\end{itemize}
	\item Trade agreements
    \begin{itemize}
			\item Now take into account impact on foreign welfare
			\item Internalize these new externalities
    \end{itemize}
\end{itemize}
\end{frame}

\begin{frame}{Role of Trade Agreements: Domestic Commitment}
\pause
\begin{itemize}[<+->]
	\item Maggi and Rodriguez-Clare (1998, 2007)
		\begin{itemize}
			\item Allow for (imperfect) capital mobility
			\item Domestic investment decisions depend on level of protection
			\item Inability to commit $\Rightarrow$ investment too high b/c importers know protection will respond
			\item Trade agreements provide commitment device
    \end{itemize}
	\item Mitra (2002)
		\begin{itemize}
			\item Here distortion is wasted resources in lobby formation
		\end{itemize}
\end{itemize}
\end{frame}

@@

\section{Tariff Caps}
\subsection{Tariff Caps}
\begin{comment}
\begin{frame}{Tariff Caps: Exogenous vs. Endogenous $\ga$}
\pause
Strong binding: an exact tariff commitment
\begin{itemize}
	\pause
	\item Must apply the precise tariff level specified in TA
\end{itemize}

\pause
\vskip.2in
Weak binding: a tariff cap
\begin{itemize}
\pause
	\item Must set tariff at or below specified level 
\end{itemize}

\pause
\vskip.4in
Perfect external enforcement vs. self-enforcement (repeated-game)

\begin{frame}{Strong Bindings}
Must set tariff at specified level
\pause
\begin{itemize}[<+->]
	\item $\ga$ exogenous (Bagwell $\&$ Staiger 2005): optimal (rigid) strong binding is the tariff that is politically optimal for the expected realization of $\ga$
  \item $\ga$ endogenous: if governments have no inherent bias toward protection, the lobbies exert no effort and are afforded no protection 
\end{itemize}
\end{frame}
\end{comment}

@@

\begin{comment}
\begin{frame}{Highlight: One-Shot Intuition}
\pause
\textbf{Legislature}
\pause
\begin{itemize}
	\item Breaks agreement if median legislator prefers $\btn$ to $\bta$
\end{itemize}

\pause
\vskip.1in
\textbf{Lobby}
\pause
\begin{itemize}[<+->]
	\item Given the $\bta$ it faces, lobby knows what $e_b$ is required to break the agreement
 	\item Lobby pays if $\bm{\pi}(\btn)$ - lobbying costs $> \pi(\bta)$
\end{itemize}

\pause
\vskip.13in
\textbf{Executives}
\pause
\begin{itemize}[<+->]
	\item Set $\bta$ to make paying $e_b$ unprofitable
		\pause
		\begin{itemize}
			\item[$\Rightarrow$] $e_b=0$, agreement remains in force
		\end{itemize}
	\item High tariffs, no lobbying, no trade disruptions
\end{itemize}
\end{frame}
\end{comment}

\begin{frame}{Repeated Game Intuition}
\pause
Legislature: break agreement if punishment not strong enough
\pause
\begin{itemize}
	\item i.e. if one period of gain from cheater's payoff is greater than T-periods of loss from trade-war
\end{itemize}

\vskip.1in
\pause
Lobby: solve for lowest effort ($\ov{e}_b$) that breaks this constraint
\pause
\begin{itemize}
	\item pay $\ov{e}_b$ if it's less than gain from $T$ periods of trade-war profits
\end{itemize}

\pause
\vskip.1in
Executives: set lowest $\bta$ that makes paying $\ov{e}_b$ unprofitable \textit{and} satisfies legislature's condition
\pause
\begin{itemize}
	\item[$\Rightarrow$] $e_b=0$, agreement remains in force
	\pause
	\item High tariffs, no lobbying, no trade disruptions
\end{itemize}
\end{frame}

@@

\begin{frame}{When the world is more complicated... (2)}
\begin{itemize}[<+->]
	\item To make escape clause work, can't use $\ga$
	\begin{itemize}[<+->]
		\item Need signal of shock that is not influenced by endogenous pressure
	\end{itemize}
	\item Can condition directly on $s$
		\begin{itemize}
			\item This seems to be what the WTO actually \textit{does}
		\end{itemize}
\end{itemize}

@@

\begin{comment}
\begin{frame}{Ex-Ante Lobbying}
Without ex-ante: gov't anticipates ex-post lobbying, chooses $\tau^{\text{TA}}$ that maximizes objective function given $e^{\text{EP}}$
  \[
	  \max_\tau W(\ga(e^{\text{EP}}),\tau) + W^*(\tau)
	\]
\pause
When ex-ante lobbying ($e^{\text{EA}}$) is possible, we could have
\pause
  \[
	  \max_\tau W(\ga(e^{\text{EA}}),\tau) + W^*(\tau)
	\]
	but the lobby would now have to pay \textit{twice}.
\end{frame}
\end{comment}

>>>>

\title[Lobbying and Legislative Uncertainty\hspace{2.95in}\insertframenumber/\inserttotalframenumber]{Lobbying and Legislative Uncertainty}
%\author[Kristy Buzard]{Kristy Buzard \\ Syracuse University and The Wallis Institute \\ kbuzard@syr.edu}
\author[Kristy Buzard]{\texorpdfstring{Kristy Buzard\newline Syracuse University and The Wallis Institute  \newline\url{kbuzard@syr.edu}}{Kristy Buzard}}
\date{April 16, 2016}
\begin{document}
\maketitle
%\insertpresentationendpage removed b/c of appendix

@@

@@

\section{Tariff Caps}
\subsection{Tariff Caps}

@@

@@

@@
